Reproducible, flexible and high throughput data extraction from primary literature: The \pkg{metaDigitise} \proglang{R} package \vskip15pt Supplementary Materials

### 2018-03-07 docs/MEE_submission/supplements.tex
supplements updates

\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb}

@@

@@

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\theequation}{S\arabic{equation}}%
\renewcommand{\thesection}{S\arabic{section}}%
\renewcommand{\thesubsection}{S\arabic{section}.\arabic{subsection}}%

@@

@@

We will now demonstrate how \fct{metaDigitise} works using figures generated from the well known iris data set. Users can install the \pkg{metaDigitise} package from GitHub as follows:

@@

Assume that the user would like to extract descriptive statistics from studies measuring sepal length or width in iris species for a fictitious project. There are a few studies that only present these data in figures. As the user reads papers found from a systematic search, they add figures with relevant data to a "FiguresToExtract" folder as follows

@@

Here, the naming of the files placed in the folder will contain the paper number, first author and the figure number to keep data uniquely associated with figures. At first there is one figure in the folder, shown in Figure \ref{fig:Anderson}. Running \fct{metaDigitise} brings up a series of prompts for the user using a main menu that provides access to a number of its features ("..." here represents the user's path to the project directory):

@@

The user simply enters in the numeric value that corresponds to what they would like to do. In this case they want to "Process new images". The user is then asked whether there are different types of plot(s) in the folder. This question is most relevant when there are lots of different figures in the folder because it will then ask the user for the type of figure as they are cycled through.

@@

\fct{metaDigitise} then asks the user whether the figure needs to be rotated or flipped. This can be needed when box plots and mean and error plots are not orientated correctly. In some cases, older papers can give slightly off angled images which can be corrected by rotating. So, in this prompt the user has three options: \code{f} for ``Flip'', \code{r} for ``rotate'' or \code{c} for ``continue''.

@@

After this, \fct{metaDigitise} will ask the user to specify the plot type. Depending on the figure, the user can specify that it is a figure containing the mean and error (\code{m}), a box plot (\code
{b}), a scatter plot (\code{s}) or a histogram (\code{h}). If the user has specified \code{d} instead of \code{s} in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.

@@

After selecting the figure type a new set of prompts will come up that will ask the user first what the y and x-axis variables are. This is useful as users can keep track of the different variables across figures and papers. Here, the user can just add this information in to the \proglang{R} console. Once complete, details on how to calibrate the x and y-axis appear, so that the relevant statistics / data can be correctly calculated. When working with a plot of mean and standard errors, the x-axis is rather useless in terms of calibration so \fct{metaDigitise} just asks the user to calibrate the y-axis.

@@

The user can just follow the instructions on screen step-by-step (instructions above have been truncated by `...' to simplify), and in the order specified. Before moving on, the user is forced to check whether or not the calibration has been set up correctly. If \code{n} is chosen because something needs to be fixed then the user can re-calibrate.

@@

Often, plots might contain multiple groups that the meta-analyst wants to extract from. \fct{metaDigitise} handles this nicely by prompting the user to enter the group first, followed by digitisation of this groups data. After digitising the first group, and having exited, \fct{metaDigitise} will ask the user whether they would like to add another group. Users can continually add groups (\code{a}), delete groups (\code{d}), edit groups (\code{e}) or finish a plot and continue to the next one (\code{f} - if another plot exists). 
%%% This needs checking - think it is now different
The number of groups are not really limited and users can just keep adding in groups to accommodate the different numbers that may be presented across figures (although it can get complicated with too many).

@@

To finish selecting points, the user can exit by clicking on the red button that appears when extracting points. The user is then asked if they want to add or delete points from that group.

@@

Users can access the \code{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the \fct{metaDigitse} function. In the \proglang{R} console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:

@@

One thing anyone with a familiarity with the iris dataset will notice is that the sample sizes for each of these species (which are n = 50 each) are quite a bit lower. This is an example of some of the challenges when extracting data from scatter plots. Often data points will overlap with each other making it impossible (without having the real data) to know whether this is a problem. However, a meta-analyst will probably realise that the sample sizes here conflict with what is reported in the paper. Hence, \pkg{metaDigitise} also provides the user with  options to input the sample sizes directly (see Editing section below), even for scatter plots and histograms where, strictly speaking, this should not be necessary. Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on standard deviation (SD) and standard error (SE). Here, the mean point estimates are nearly exactly the same as the true values, but the SD's are slightly over-estimated:

@@

Users can add additional figures as new papers with relevant information are found. Each figure should be in its own file with unique naming, even if a single paper has multiple figures for extraction. For example, another paper on different populations (and one new species) of iris contained two additional figures where important data could be extracted. These figures can simply be named accordingly and added directly to the same extraction folder:

@@

The user has already processed one figure (001\_Anderson\_1935\_Fig1.png). We can tell this because the caldat folder has digitised data in it (caldat/001\_Anderson\_1935\_Fig1). Now the user has two new figures that have not yet been digitised. This example will nicely demonstrate how users can easily pick up from where they left off and how all previous data gets re-integrated. It will also demonstrate how different plot types are handled. All we have to do to begin, is again, provide the directory where all the figures are located:

@@

The user gets the same set of prompts and simply chooses option one. This will permit users to digitise new figures, and will integrate previously completed digitisations along with newly digitised data together at the end of the session, or when the user decides to quit. This time, 001\_Anderson\_1935\_Fig1.png is ignored and the new plots cycle on screen; first 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png. Since there are a few different figure types, the user answers the first question in the \proglang{R} console as "d":

@@

Here, the user specifies the new plot type as \code{m} for 002\_Doe\_2013\_Fig1.png because the user has a plot of the mean and error of sepal length for each of the three species. The user is then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:

@@

Again, \fct{metaDigitise} will simply guide the user through digitising each of these figures describing to them exactly what needs to be done. At any point if mistakes are made the user can choose relevant options to edit or correct things before ending the figure. This process continues for each plot so long as the user would like to continue and after completing a single plot the user is always prompted as follows:

@@

This continues until users have completed all non-digitised figures in the folder, at which point \fct{metaDigitise} concatenates the new data with previously digitised data in the object:

@@

If the user chooses "Import existing data", they have the option of either 1) importing data from all digitised images or 2) importing data from a single image that has been digitised. If 2, then a list of files are provided to the user that they can select. Editing existing data allows users to easily re-plot or edit information or digitisations that have previously be done for any plot. This is accomplished by guiding the user through a new set of options:

@@

If the user is unsure about the name of the specific figure they need to edit or simply want to just check the digitisations of figures they can choose "Cycle through images", which will bring up each figure, one by one, overlaying the calibrations, group names (if they exist), sample sizes (if they were entered) and the selected points. The user will then be given the choice to edit individual images. Alternatively, choosing option 2, will bring up a list of the completed files in the folder and the specific file can be chosen, at which point it will be replotted. Either of these options will cycle through a number of questions asking the user what they would like to edit:

@@

\begin{CodeOutput}
Edit rotation? If yes, then the whole extraction will be redone (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Change plot type? If yes, then the whole extraction will be redone (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Variable entered as: 
\end{CodeOutput}
\begin{CodeInput}
R> Sepal length
\end{CodeInput}
\begin{CodeOutput}
Rename Variables (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Edit calibration? (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Re-extract data (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> y
\end{CodeInput}
\begin{CodeOutput}
Change group identifier? (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Add group, Delete group or Finish plot? (a/d/f) 
\end{CodeOutput}
\begin{CodeInput}
R> d
\end{CodeInput}
\begin{CodeOutput}
1: setosa
2: versicolor
3: viriginica
Selection: 
\end{CodeOutput}
\begin{CodeInput}
R> 2
\end{CodeInput}
\begin{CodeOutput}
Add group, Delete group or Finish plot? (a/d/f) 
\end{CodeOutput}
\begin{CodeInput}
R> a
\end{CodeInput}

@@

While \fct{metaDigitise} provides users with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the processed digitised data from scatter plots (i.e. calibrated points). This is very easy to do my changing the default \code{summary} argument from TRUE to FALSE in \fct{metaDigitise}. Instead of providing the user with summary statistics it will return a list containing four slots for each of the figure types (mean error, box plot, histogram and scatter plots). An example of a data object returned from digitising figures is as follows:

@@

\subsection{Adding sample sizes to previous Digitisations}
%Including addition of N later
In many cases important information, such as sample sizes, may not be readily available or clear when digitising figures. In these circumstances users will have answered `no' to the question about whether they have sample sizes or not while digitising. To expedite finding and adding in these sample sizes to do the necessary calculations (if for example a figure presented 95\% CI's or standard errors), \fct{metaDigitise} has s specific edit option that allows users to enter in previously omitted sample sizes. It works by first identifying the missing sample sizes in the digitised output, re-plotting the relevant figure and then prompting the user to enter the sample sizes for the relevant groups in the figure, one by one. As an example, assume that we were missing sample sizes for two groups in 002\_Doe\_2013\_Fig1.png:

@@

Here, we can see that we are missing the sample sizes for setosa and viriginica, and as a result, sd is not calculated because \fct{metaDigitise} needs this information to make the calculation. If the user found this information after contacting the authors for clarification then they can add these back in as follows:

@@

\fct{metaDigitise} will replot the figure after this and list, only the groups missing data, for which the user can then update the data. This is then re-integrated back into the data automatically and the sd calculated.

@@

\subsection{Figure Rotation and Adjustment}
Figures may have been extracted from old publications, for example from scanned images, and so are not perfectly orientated on the image. This will make the calibration of the points in the figure from the image problematic. \fct{metaDigitise} allows users to rotate the image. By clicking two points on the x-axis, metaDigitse calculates the angle needed to rotate the image so the x-axis is horizontal, and rotates it. (Figure \ref{fig:rotate}A,B)

\begin{figure}[!b] 
\centering
 \includegraphics[width=0.75\textwidth]{fig_rotate.pdf} 
 \caption{Figure rotation. A) and B) show how non-aligned images can be realigned through user defined rotation. C) and D) show how figures can be re-orientated so as to aid data input.}
\label{fig:rotate}
\end{figure}

Furthermore, some figures, including mean and error, boxplots or histograms, may be presented with horizontal bars. \fct{metaDigitise} assumes that the bars are vertical, but allows the user to flip the image so that the bars are vertical if provided horizontally (Figure \ref{fig:rotate}C,D).

@@

\section{Software Validation: Details of Simulation}

\subsection{Inter-observer variability in digitisations}
In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups($n=10$ within groups). Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures. Fourteen independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order. Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux). Digitisers varied in their level of experience, from people with experience of meta-analyses to those without any science background. We asked users to digitise all 14 figures and collected the mean, standard deviation and correlation coefficient ($r$, for scatter plots) generated by \fct{metaDigitise} for every plot digitised ($n=28$ per digitiser per metric, $n=4$ for $r$).

@@

Across plot types we found no evidence for any inter-observer variability in digitisations for the mean (ICC = 0, 95\% CI = 0 to 0.029, \textit{p} = 1), standard deviation (ICC = 0, 95\% CI = 0 to 0.033, \textit{p} = 0.5) or correlation coefficient (ICC = 0.053, 95\% CI = 0 to 0.296, \textit{p} = 0.377). There were was little bias between digitised and true values, on average 1.63\% (mean = 0.02\%, SD = 4.9\%, \textit{r} = -0.03\%) and overall there were only small absolute differences between digitised and true values, deviating, on average 2.18\% (mean = 0.40\%, SD = 5.81\%, \textit{r} = 0.33\%) across all three summary statistics. SD estimates from digitisations are clearly more prone to error than means or correlation coefficients. If the mean absolute difference is calculated for each plot type, we can see that this effect is driven mainly by extraction from box plots and histograms (\% difference; box plot: 15.805, histogram: 5.210, mean/error: 1.500, scatter plot: 0.433). SD estimation from box plot summary statistics is known to be more error prone, especially at small sample sizes \citep{Wan2014}.

\subsection{Testing the accuracy of digitisations}
To test how accurate \pkg{metaDigitise} is at matching clicked points to their true values, we generated four random scatterplots, each with 20 data points, and digitised these with \fct{metaDigitise}. This was done by one digitiser (J.L.P.), as there is no detectable between user variation. Data digitised using \pkg{metaDigitise} was essentially perfectly correlated with the true simulated data for both the \textit{x}-variable (Pearson's correlation; \textit{r} = 0.9999915, \textit{t} = 2137.4, \textit{df} = 78, $p < 0.001$) and \textit{y}-variable (\textit{r} = 0.9999892, \textit{t} = 1897.8, \textit{df} = 78, $p < 0.001$).
\section{Comparison with existing software}
\begin{sidewaystable}[ph!]
{\centering
\begin{tabular}{lccccccc}
\hline
Function                  & metaDigitise & GraphClick$^1$ & DataThief$^2$  & DigitizeIt$^3$ & WebPlotDigitizer$^4$ & metagear$^5$ & digitize$^6$\\
\hline
Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & \checkmark$^7$ & \checkmark \\
Mean/error plots      & \checkmark   & \checkmark & \checkmark & $\times$   & $\times$       & \checkmark$^7$ & $\times$    \\
Boxplots                  & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Histograms                & \checkmark   & $\times$   & $\times$   & $\times$   & \checkmark$^7$ & $\times$       & $\times$    \\
Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Groups                    & \checkmark   & \checkmark & $\times$   & \checkmark & \checkmark     & $\times$       & $\times$    \\
Entry of metadata         & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Summarising data          & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Multiple image processing & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
Automated point detection & $\times$     & \checkmark &     $\times$      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
Line extraction           & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Zoom                      & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Log axis                  & $\times$   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Dates                     & $\times$   & $\times$   & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
Asymmetric error bars     & $\times$     & $\times$   & \checkmark & $\times$   & $\times$       & $\times$       & $\times$    \\
Freeware                  & \checkmark$^{10}$ & \checkmark$^{11}$ & \checkmark$^{11}$ & $\times$$^{11}$ & \checkmark$^{11}$ & \checkmark$^{10}$ & \checkmark$^{10}$\\
\hline

@@

\end{tabular}
}
{\footnotesize
\\
$^1$ \citet{GraphClick}
$^2$ \citet{DataThief}
$^3$ \citet{DigitizeIt}
$^4$ \citet{WebPlotDigitizer}
$^5$ \citet{Lajeunesse2016}
$^6$ \citet{Poisot2011}
\\$^7$ Only automated, no manual extraction.
\\$^8$ Or handles rotated graphs. 
\\$^9$ Allows saving, re-plotting and editing of data extraction.
\\$^{10}$ \proglang{R} package.
\\$^{11}$ Standalone software.
}

\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}

\end{sidewaystable}

>>>>

\usepackage{amssymb,rotating,natbib,graphicx,fancyvrb,booktabs}

@@

\usepackage[colorlinks=true,linkcolor=blue, citecolor=blue, urlcolor=blue, linktoc=page]{hyperref}

@@

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\theequation}{S\arabic{equation}}%
\renewcommand{\thesection}{S\arabic{section}}%
\renewcommand{\thesubsection}{S\arabic{section}.\arabic{subsection}}%

@@

\tableofcontents

@@

We can demonstrate how \fct{metaDigitise} works using figures generated from the well known iris data set. \pkg{metaDigitise} can installed from GitHub as follows:

@@

Assume you would like to extract descriptive statistics from studies measuring sepal length or width in iris species for a fictitious project. There are a few studies that only present these data in figures. As you read papers found from a systematic search, you add figures with relevant data to a "FiguresToExtract" folder as follows

@@

Here, the naming of the files placed in the folder will contain the paper number, first author and the figure number to keep data uniquely associated with figures. At first there is one figure in the folder, shown in Figure \ref{fig:Anderson}. Running \fct{metaDigitise} brings up a series of prompts using a main menu that provides access to a number of its features ("..." here represents the user's path to the project directory):

@@

You then simply enter in the numeric value that corresponds to what you would like to do. In this case we want to "Process new images". You are then asked whether there are different types of plot(s) in the folder. This question is most relevant when there are lots of different figures in the folder because it will then ask for the type of figure as they are cycled through.

@@

\fct{metaDigitise} then asks whether the figure needs to be rotated or flipped. This can be needed when box plots and mean and error plots are not orientated correctly. In some cases, older papers can give slightly off angled images which can be corrected by rotating. In this prompt the user has three options: \code{f} for ``Flip'', \code{r} for ``rotate'' or \code{c} for ``continue''.

@@

After this, \fct{metaDigitise} will ask you to specify the plot type. Depending on the figure, you can specify that it is a figure containing the mean and error (\code{m}), a box plot (\code{b}), a scatter plot (\code{s}) or a histogram (\code{h}). If you has specified \code{d} instead of \code{s} in response to the question about whether the plot types are the same or different, this question will pop up for each plot, but will only be asked once if plots are all the same.

@@

After selecting the figure type a new set of prompts will come up that will ask what the y and x-axis variables are. This is useful as users can keep track of the different variables across figures and papers. Here, you can just add this information in to the \proglang{R} console. Once complete, details on how to calibrate the x and y-axis appear, so that the relevant statistics / data can be correctly calculated. When working with a plot of mean and errors, the x-axis is rather useless in terms of calibration so \fct{metaDigitise} just asks you to calibrate the y-axis.

@@

You can just follow the instructions on screen step-by-step (instructions above have been truncated by `...' to simplify), and in the order specified. Before moving on, you are forced to check whether or not the calibration has been set up correctly. If \code{n} is chosen because something needs to be fixed then you can re-calibrate.

@@

Often, plots might contain multiple groups that a meta-analyst wants to extract from. \fct{metaDigitise} handles this nicely by prompting you to enter the group first, followed by digitisation of this groups data. After digitising the first group, \fct{metaDigitise} will ask whether you would like to add another group. You can continually add groups (\code{a}), delete groups (\code{d}), edit groups (\code{e}) or finish a plot and continue to the next one (\code{f} - if another plot exists). 
%%% This needs checking - think it is now different
The number of groups are not really limited and you can just keep adding in groups to accommodate the different numbers that may be presented across figures (although it can get complicated with too many).

@@

To finish selecting points, you can exit by clicking on the red button that appears when extracting points. You are then asked if you want to add or delete points from that group.

@@

You can access the \code{metaDigitise} object created (001\_Anderson\_1935\_Fig1) at any time using the \fct{metaDigitse} function. In the \proglang{R} console, the summarised data for the digitised figure can be printed on screen or even written to a .csv file:

@@

One thing anyone with a familiarity with the iris dataset will notice is that the sample sizes for each of these species (which are n = 50 each) are quite a bit lower. This is an example of some of the challenges when extracting data from scatter plots. Often data points will overlap with each other making it impossible (without having the real data) to know whether this is a problem. However, a meta-analyst will probably realise that the sample sizes here conflict with what is reported in the paper. Hence, \pkg{metaDigitise} also provides you with  options to input the sample sizes directly, even for scatter plots and histograms where, strictly speaking, this should not be necessary. Nonetheless, it is important to recognise the impact that overlapping points can have on summary statistics, particularly its effects on sampling variance. Here, the mean point estimates are nearly exactly the same as the true values, but the SD's are slightly over-estimated:

@@

You can add additional figures to the directory as new papers with relevant information are found. Each figure should be in its own file with unique naming, even if a single paper has multiple figures for extraction. For example, another paper on different populations (and one new species) of iris contained two additional figures where important data could be extracted. These figures can simply be named accordingly and added directly to the same extraction folder:

@@

We have already processed one figure (001\_Anderson\_1935\_Fig1.png). We can tell this because the caldat folder has digitised data in it (caldat/001\_Anderson\_1935\_Fig1). Now there are two new figures that have not yet been digitised. This example will also nicely demonstrate how you can easily pick up from where you left off and how previous data gets re-integrated. It will also demonstrate how different plot types are handled. All we have to do to begin, is again, provide the directory where all the figures are located:

@@

You get the same set of prompts and simply chooses option one. This will permit users to digitise new figures, and will integrate previously completed digitisations along with newly digitised data together at the end of the session, or when the you decides to quit. This time, 001\_Anderson\_1935\_Fig1.png is ignored and the new plots cycle on screen; first 002\_Doe\_2013\_Fig1.png and then 002\_Doe\_2013\_Fig3.png. Since there are a few different figure types, you answer the first question in the \proglang{R} console as "d":

@@

Here, you specify the new plot type as \code{m} for 002\_Doe\_2013\_Fig1.png because this a plot of the mean and error of sepal length for each of the three species. The user is then prompted a bit differently from our scatter plot as the x-axis is not needed for calibration:

@@

Again, \fct{metaDigitise} will simply guide you through digitising each of these figures describing to you exactly what needs to be done. At any point if mistakes are made you can choose relevant options to edit or correct things before ending the figure. This process continues for each plot so long as the you would like to continue and after completing a single plot you are always prompted as follows:

@@

This continues until you have completed all non-digitised figures in the folder, at which point \fct{metaDigitise} concatenates the new data with previously digitised data in the object:

@@

If we choose "Import existing data", you have the option of either 1) importing data from all digitised images or 2) importing data from a single image that has been digitised. If 2, then a list of files are provided for you to select. Editing existing data allows you to easily re-plot or edit information or digitisations that have previously be done for any plot. This is accomplished by guiding you through a new set of options:

@@

If you is unsure about the name of the specific figure you need to edit or simply want to just check the digitisations of figures they can choose "Cycle through images", which will bring up each figure, one by one, overlaying the calibrations, group names (if they exist), sample sizes (if they were entered) and the selected points. The user will then be given the choice to edit individual images. Alternatively, choosing option 2, will bring up a list of the completed files in the folder and the specific file can be chosen, at which point it will be replotted. Either of these options will cycle through a number of questions asking you what they would like to edit:

@@

\begin{CodeOutput}
Edit rotation? If yes, then the whole extraction will be redone (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Change plot type? If yes, then the whole extraction will be redone (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Variable entered as: 
Sepal length
Rename Variables (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Edit calibration? (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Re-extract data (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> y
\end{CodeInput}
\begin{CodeOutput}
Change group identifier? (y/n) 
\end{CodeOutput}
\begin{CodeInput}
R> n
\end{CodeInput}
\begin{CodeOutput}
Add group, Delete group or Finish plot? (a/d/f) 
\end{CodeOutput}
\begin{CodeInput}
R> d
\end{CodeInput}
\begin{CodeOutput}
1: setosa
2: versicolor
3: viriginica
Selection: 
\end{CodeOutput}
\begin{CodeInput}
R> 2
\end{CodeInput}
\begin{CodeOutput}
Add group, Delete group or Finish plot? (a/d/f) 
\end{CodeOutput}
\begin{CodeInput}
R> a
\end{CodeInput}

@@

While \fct{metaDigitise} provides you with the summary statistics by default, for all plot types, in many cases the user may actually be interested in obtaining the processed digitised data from scatter plots (i.e. calibrated points). This is very easy to do by changing the default \code{summary} argument from TRUE to FALSE in \fct{metaDigitise}. Instead of providing you with summary statistics it will return a list containing four slots for each of the figure types (mean error, box plot, histogram and scatter plots). An example of a data object returned from digitising figures is as follows:

@@

Additionally there is another function, \fct{import\_metaDigitise}, that imports all digitised data from a directory, as either summarised or processed data. For example,

\begin{CodeChunk}
\begin{CodeInput}
>R import_metaDigitise(".../FiguresToExtract", summary=TRUE)
\end{CodeInput}
{\scriptsize
\begin{CodeOutput}
                  filename    group_id         variable   mean  error error_type n    r    sd   plot_type
001_Anderson_1935_Fig1.png      setosa  Sepal width (mm)  3.42  0.40  sd        39  0.75  0.40  scatterplot
001_Anderson_1935_Fig1.png      setosa  Sepal length (mm) 5.00  0.38  sd        39  0.75  0.38  scatterplot
001_Anderson_1935_Fig1.png  versicolor  Sepal width (mm)  2.77  0.32  sd        44  0.52  0.32  scatterplot
001_Anderson_1935_Fig1.png  versicolor  Sepal length (mm) 5.95  0.53  sd        44  0.52  0.53  scatterplot
     002_Doe_2013_Fig1.png      setosa  Sepal length      5.00  0.11  se        50  NA    0.78  mean_error
     002_Doe_2013_Fig1.png  viriginica  Sepal length      6.59  0.18  se        50  NA    1.26  mean_error
     002_Doe_2013_Fig1.png  versicolor  Sepal length      5.94  0.14  se        50  NA    1.01  mean_error
     003_Doe_2013_Fig3.png      catana  Sepal length      4.95  0.36  sd        50  NA    0.36  histogram
\end{CodeOutput}
}
\end{CodeChunk}

This function is useful in scripts the extracted data simply needs to be imported, and where the interactive element of \fct{metaDigitise} becomes problematic.

\subsection{Adding sample sizes to previous Digitisations}
%Including addition of N later
In many cases important information, such as sample sizes, may not be readily available or clear when digitising figures. In these circumstances you would have answered `no' to the question about whether you have sample sizes or not while digitising. To expedite finding and adding in these sample sizes to do the necessary calculations (if for example a figure presented 95\% CI's or standard errors), \fct{metaDigitise} has s specific edit option that allows users to enter in previously omitted sample sizes. It works by first identifying the missing sample sizes in the digitised output, re-plotting the relevant figure and then prompting you to enter the sample sizes for the relevant groups in the figure, one by one. As an example, assume that we were missing sample sizes for two groups in 002\_Doe\_2013\_Fig1.png:

@@

Here, we can see that we are missing the sample sizes for setosa and viriginica, and as a result, sd is not calculated because \fct{metaDigitise} needs this information to make the calculation. If, for examnple, you found this information after contacting the authors for clarification then you can add these in as follows:

@@

\fct{metaDigitise} will replot the figure after this and list, only the groups missing data, for which you can then update the data. This is then re-integrated back into the data automatically and the sd calculated.

@@

@@

\section{Software Validation: Details of Simulations}

\subsection{Inter-observer variability in digitisations}
In order to evaluate the consistency of digitisation using \pkg{metaDigitise} between users, we simulated a dataset of two variables with two groups ($n=10$ within groups). Each variable was plotted twice for each plot type (figures were modified slightly to give users a sense that they were digitising new data) generating a total of 14 figures. Fourteen independent digitisers (including the authors) were provided with a directory with all 14 figures in a randomised order. Digitisers ran \pkg{metaDigistise} on their own computers, across different operating systems (including Mac, Windows and Linux). Digitisers varied in their level of experience, from people with experience of meta-analyses to those without any science background. We asked users to digitise all 14 figures and collected the mean, standard deviation and correlation coefficient ($r$, for scatter plots) generated by \fct{metaDigitise} for every plot digitised ($n=28$ per digitiser per metric, $n=4$ for $r$).

@@

\subsection{Testing the accuracy of digitisations}
To test how accurate \pkg{metaDigitise} is at matching clicked points to their true values, we generated four random scatterplots, each with 20 data points, and digitised these with \fct{metaDigitise}. This was done by one digitiser (J.L.P.), as there is no detectable between user variation.

\section{Comparison with existing software}
\begin{table}[h!]
\centering
\rotatebox{90}{%
\begin{minipage}{1.35\textwidth}
\small
\begin{tabular}{lccccccc}
\hline
Function                  & metaDigitise & GraphClick$^1$ & DataThief$^2$  & DigitizeIt$^3$ & WebPlotDigitizer$^4$ & metagear$^5$ & digitize$^6$\\
\hline
Scatterplots              & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & \checkmark$^7$ & \checkmark \\
Mean/error plots      & \checkmark   & \checkmark & \checkmark & $\times$   & $\times$       & \checkmark$^7$ & $\times$    \\
Boxplots                  & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Histograms                & \checkmark   & $\times$   & $\times$   & $\times$   & \checkmark$^7$ & $\times$       & $\times$    \\
Graph rotation$^8$        & \checkmark   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Groups                    & \checkmark   & \checkmark & $\times$   & \checkmark & \checkmark     & $\times$       & $\times$    \\
Entry of metadata         & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Summarising data          & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Multiple image processing & \checkmark   & $\times$   & $\times$   & $\times$   & $\times$       & $\times$       & $\times$    \\
Reproducable$^9$  & \checkmark   & \checkmark & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
Automated point detection & $\times$     & \checkmark &     $\times$      & \checkmark & \checkmark     & \checkmark     & $\times$    \\
Line extraction           & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Zoom                      & $\times$     & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Log axis                  & $\times$   & \checkmark & \checkmark & \checkmark & \checkmark     & $\times$       & $\times$    \\
Dates                     & $\times$   & $\times$   & \checkmark & $\times$   & \checkmark     & $\times$       & $\times$    \\
Asymmetric error bars     & $\times$     & $\times$   & \checkmark & $\times$   & $\times$       & $\times$       & $\times$    \\
Freeware                  & \checkmark$^{10}$ & \checkmark$^{11}$ & \checkmark$^{11}$ & $\times$$^{11}$ & \checkmark$^{11}$ & \checkmark$^{10}$ & \checkmark$^{10}$\\
\hline

@@

\end{tabular}
\\
$^1$ \citet{GraphClick}
$^2$ \citet{DataThief}
$^3$ \citet{DigitizeIt}
$^4$ \citet{WebPlotDigitizer}
$^5$ \citet{Lajeunesse2016}
$^6$ \citet{Poisot2011}
\\$^7$ Only automated, no manual extraction.
\\$^8$ Or handles rotated graphs. 
\\$^9$ Allows saving, re-plotting and editing of data extraction.
\\$^{10}$ \proglang{R} package.
\\$^{11}$ Standalone software.
\caption{\label{tab:comparison} Comparison of functionality between different digitisation softwares.}
\end{minipage}
}

\end{table}
